---
title: "Introduction to clustermole"
---

```{r, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
# reduce the minimum number of characters for the tibble column titles (default: 15)
options(pillar.min_title_chars = 10)
```

## Overview

The clustermole R package was developed to assist with identifying candidate cell type labels for single-cell RNA-seq (scRNA-seq) clusters.
The package includes three primary features:

- Retrieve human and mouse cell type signatures with `clustermole_markers()`.
- Compare a set of marker genes with the signatures using `clustermole_overlaps()`.
- Score signature enrichment across an expression matrix without selecting marker genes using `clustermole_enrichment()`.

All three functions return tables that can be easily filtered, summarized, and used in other workflows.

The reference meta-database combines and cleans cell type markers from multiple sources.
See [database details](db.html) for an overview of the contents and the curation process.
Sources differ in their studies, tissues, curation methods, and selection criteria.
A marker may identify a broad cell class, a subtype, or a cell state.
Comparing results across sources can reveal both recurring candidates and alternative labels worth investigating that a single source may miss.

For examples of applications across species, tissues, and technologies, see [publications using clustermole](pubs.html).

## Setup

clustermole is available from [CRAN](https://cran.r-project.org/package=clustermole).

```{r install-package, eval=FALSE}
install.packages("clustermole")
```

Load clustermole.

```{r load-package, message=FALSE, warning=FALSE}
library(clustermole)
```

## Cell type markers

You can use clustermole as a simple database and get a data frame of all cell type markers.
The `species` argument selects human or mouse gene symbols.
It does not restrict signatures to the selected species.

```{r markers}
markers <- clustermole_markers(species = "hs")
markers
```

The output data frame is in an R-friendly tidy/long format with one gene-to-signature mapping per row.
Each row contains a gene and a cell type associated with it.
The `gene` column is the gene symbol and the `celltype_full` column contains the full cell type string (cell type, organ/tissue, species, and source).
The `species` column contains the source species which can differ from the selected gene symbol species.

Many tools that work with gene sets require input as a list.
You can use `split()` to convert the data frame to a list using `gene` as the value and `celltype_full` as the grouping variable.

```{r celltypes-list}
markers <- dplyr::distinct(markers, celltype_full, gene)
markers_list <- split(x = markers$gene, f = markers$celltype_full)
```

## Cell types based on marker genes

If you have a vector of genes, such as cluster markers, you can compare them to known cell type markers to see if they overlap any of the known cell type markers (over-representation analysis).

```{r overlaps, eval=FALSE}
my_overlaps <- clustermole_overlaps(genes = my_genes, species = "hs")
```

## Cell types based on an expression matrix

If you have expression values, such as average expression for each cluster, you can perform cell type enrichment based on the full gene expression matrix (log-transformed CPM/TPM/FPKM values).
The underlying enrichment method can be changed using the `method` parameter.

```{r enrichment, eval=FALSE}
my_enrichment <- clustermole_enrichment(expr_mat = my_expr_mat, species = "hs")
```

Bioconductor packages GSVA, GSEABase, and singscore need to be installed to use `clustermole_enrichment()`.

```{r install-bioc-packages, eval=FALSE}
BiocManager::install(c("GSVA", "GSEABase", "singscore"))
```

See the [bone marrow tutorial](example-bm-seurat.html) for examples based on a scRNA-seq dataset.
