---
title: "Database details"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
# reduce the minimum number of characters for the tibble column titles (default: 15)
options(pillar.min_title_chars = 10)
```

The clustermole meta-database collects gene signatures from public marker databases and cell atlases.
The same cell type can have several signatures that can differ by organ, species, and study.

The database includes data from the following sources:

- CellMarker ([Hu et al. *Nucleic Acids Research* 2023](https://doi.org/10.1093/nar/gkac947))
- CellMatch ([Shao et al. *iScience* 2020](https://doi.org/10.1016/j.isci.2020.100882))
- Cell Taxonomy ([Jiang et al. *Nucleic Acids Research* 2023](https://doi.org/10.1093/nar/gkac816))
- DISCO ([Li et al. *Nucleic Acids Research* 2022](https://doi.org/10.1093/nar/gkab1020))
- HPA ([Karlsson et al. *Science Advances* 2021](https://doi.org/10.1126/sciadv.abh2169))
- MSigDB ([Liberzon et al. *Bioinformatics* 2011](https://doi.org/10.1093/bioinformatics/btr260))
- PanglaoDB ([Franzen et al. *Database* 2019](https://doi.org/10.1093/database/baz046))
- SaVanT ([Lopez et al. *BMC Genomics* 2017](https://doi.org/10.1186/s12864-017-4167-7))
- ScType ([Ianevski et al. *Nature Communications* 2022](https://doi.org/10.1038/s41467-022-28803-w))
- TISSUES ([Palasca et al. *Database* 2018](https://doi.org/10.1093/database/bay003))
- xCell ([Aran et al. *Genome Biology* 2017](https://doi.org/10.1186/s13059-017-1349-1))

Many of these are static, but some are updated periodically.

The raw source data is processed to make the structure and content consistent across databases.
Database preparation steps include:

- Converting recognized human and mouse gene aliases to canonical symbols.
- Checking that the species labels are supported by the provided genes.
- Removing signatures with fewer than five or more than 1,000 genes.
- Mapping genes to orthologs in the other species.

## Database data frame

The summaries below describe the database based on the actual package contents.

```{r load-libraries, message=FALSE, warning=FALSE}
library(clustermole)
library(dplyr)
```

Retrieve a data frame of all cell type markers in the database.
The `species` argument selects the appropriate gene symbols and does not restrict signatures to the selected species.

```{r markers}
markers <- clustermole_markers(species = "hs")
markers
```

The output data frame is in an R-friendly tidy/long format with one gene-to-signature mapping per row.

## Total cell types

Check the total number of available cell type signatures.
Each `celltype_full` value uniquely identifies a signature (cell type, organ/tissue, species, and source).

```{r celltypes-length}
length(unique(markers$celltype_full))
```

## Cell types per source

Check the number of cell type signatures from each source database.

```{r count-db}
distinct(markers, db, celltype_full) |> count(db)
```

## Cell types per species

Check the number of cell type signatures per original source species.
Not all sources provide information about the species, so it can be unknown or blank.

```{r count-species}
distinct(markers, species, celltype_full) |> count(species)
```

## Cell types per organ

Check the number of available cell type signatures per organ or tissue.
This label is not standardized and is not always available.

```{r count-organ}
distinct(markers, celltype_full, organ) |> count(organ, sort = TRUE)
```

## Package version

Check the package version since the database contents can change.

```{r package-version}
packageVersion("clustermole")
```
