---
title: "Database statistics"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

We will load `clustermole` along with `dplyr` to help with summarizing the data.

```{r setup, message=FALSE}
library(clustermole)
library(dplyr)
```

You can use `clustermole` as just a database to get a data frame of all cell type markers.

```{r markers}
markers = clustermole_markers()
markers
```

Each row contains a gene and a cell type associated with it.
The `gene` column is the gene symbol (human or mouse) and the `celltype_full` column contains the detailed cell type string including the species and the original database.

Check the total number of the available cell types.

```{r celltypes-length}
markers %>% distinct(celltype_full) %>% nrow()
```

Check the source databases.

```{r count-db}
markers %>% distinct(celltype_full, db) %>% count(db)
```

Check the number of available cell types per species (not available for every cell type).

```{r count-species}
markers %>% distinct(celltype_full, species) %>% count(species)
```

Check the number of available cell types per organ (not available for every cell type).

```{r count-organ}
markers %>% distinct(celltype_full, organ) %>% count(organ, sort = TRUE)
```

Check the package version.

```{r package-version}
packageVersion("clustermole")
```

